Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-01 ms.tex
rest

Assume that the outflow power is proportional to the SMBH accretion rate, $L_{\rm j} =
\epsilon_{\rm j} \dot{M}c^{2}$, where $\epsilon_{\rm j} < 1$ is a jet efficiency factor.  Further assume that this energy is deposited as heat uniformly interior to a radius $r_{\rm heat}$ and volume $V_{\rm heat} \propto r_{\rm
heat}^{3}$.  The resulting volumetric heating rate $\dot{e} = \epsilon_{\rm
j}\dot{M}c^{2}/V_{\rm heat}$ near the stagnation radius results in a heating parameter given by
\begin{eqnarray} v_{w}^{\bullet} &\approx& \left(\frac{2t_{\rm
h}\epsilon_{\rm j}\dot{M}c^{2}}{V_{\rm heat}\eta
\rho_{\star}}\right)^{1/2} \sim \left(\frac{\epsilon_{\rm
j}\dot{M}c^{2}t_{\rm h}}{\eta M_{\rm enc}}\right)^{1/2} \nonumber \\
&\approx& 300\,{\rm km\,s^{-1}}\,\left(\frac{\epsilon_{\rm j}}{10^{-6}}\right)^{1/2}\left(\frac{r_{\rm s}}{r_{\rm
heat}}\right)^{1-\Gamma/2},
\label{eq:vjet}
\end{eqnarray} 
where we have used the facts that $\dot{M} = \eta M_{\star}|_{r_{\rm s}}/t_{\rm h}$ and, for radii inside the Nuker break radius, $M_{\star}(r) \propto
M_{\bullet}(r/r_{\rm soi})^{2-\Gamma}$ (eq.~[\ref{eq:rhostar}]).  If the bulk of the outflow energy is released near the stagnation radius, then even a small heating efficiency $\epsilon_{\rm j} \gtrsim
10^{-5}$ is sufficient for $v_{w}^{\bullet}$ to exceed other sources of non-accretion powered heating.  However, if this energy is deposited over much larger physical scales comparable to the size of the galaxy, i.e. $r_{\rm heat} \gtrsim $ 10 kpc $\sim 10^{4}r_{\rm s}$, then kinetic feedback is unimportant, even for a powerful jet with $\epsilon_{\rm j} \sim 1$.

@@

\subsection{Black Hole Accretion Rates}

@@

\subsection{Nuclear X-ray Luminosities}
\label{sec:mdot}

@@

\subsection{Local Growth Time of SMBHs }

@@

\subsection{TDE Jets}

@@

\subsection{Jet Feedback}

We also note
here that the kinetic heating rate is proportional to the SMBH mass to
a positive power, $v_{w}^{\bullet} \propto M_{\bullet}^{1(0.5)}$ for
core(cusp) galaxies, respectively. This dependence could have
important implications for the dependence of the average nuclear X-ray
luminosity of quiescent galaxies on the SMBH mass
($\S\ref{sec:mdot}$).

\subsection{Jet Power Correlation}

Accretion rates onto SMBHs can in principle be estimated directly by
assuming Bondi accretion and using high resolution X-ray observations
to estimate the gas temperature and density near the Bondi radius.
This technique has led to the intriguing conclusion that the Bondi
accretion power correlates with the AGN jet power (e.g.,
\citealt{AllenDunn+:2006a}; \citealt{Russell+13};
\citealt{FujitaKawakatu+:2014a}), the latter estimated from the energy
required to inflate the observed cavities within the X-ray emitting
gas.  In practice is it almost never possible to resolve the Bondi
radius, therefore requiring all such studies to extrapolate the
observed gas temperature and density profiles to smaller radial
scales.

@@

>>>>

Assume that the outflow power is proportional to the SMBH accretion rate, $L_{\rm j} =
\epsilon_{\rm j} \dot{M}c^{2}$, where $\epsilon_{\rm j} < 1$ is a jet efficiency factor.  Further assume that this energy is deposited as heat uniformly interior to a radius $r_{\rm heat}$ and volume $V_{\rm heat} \propto r_{\rm
heat}^{3}$.  The resulting volumetric heating rate $\dot{e} = \epsilon_{\rm
j}\dot{M}c^{2}/V_{\rm heat}$ near the stagnation radius results in a heating parameter given by
\begin{eqnarray} v_{w}^{\bullet} &\approx& \left(\frac{2t_{\rm
h}\epsilon_{\rm j}\dot{M}c^{2}}{V_{\rm heat}\eta
\rho_{\star}|_{r_{s}}}\right)^{1/2} \sim \left(\frac{\epsilon_{\rm
j}\dot{M}c^{2}t_{\rm h}}{\eta M_{\rm enc}}\right)^{1/2} \nonumber \\
&\approx& 300\,{\rm km\,s^{-1}}\,\left(\frac{\epsilon_{\rm j}}{10^{-6}}\right)^{1/2}\left(\frac{r_{\rm s}}{r_{\rm
heat}}\right)^{1-\Gamma/2},
\label{eq:vjet}
\end{eqnarray} 
where we have used the facts that $\dot{M} = \eta M_{\star}|_{r_{\rm s}}/t_{\rm h}$ and, for radii inside the Nuker break radius, $M_{\star}(r) \propto
M_{\bullet}(r/r_{\rm soi})^{2-\Gamma}$ (eq.~[\ref{eq:rhostar}]).  If the bulk of the outflow energy is released near the stagnation radius, then even a small heating efficiency $\epsilon_{\rm j} \gtrsim
10^{-5}$ is sufficient for $v_{w}^{\bullet}$ to exceed other sources of non-accretion powered heating.  However, if this energy is deposited over much larger physical scales comparable to the size of the galaxy, i.e. $r_{\rm heat} \gtrsim $ 10 kpc $\sim 10^{4}r_{\rm s}$, then kinetic feedback is unimportant, even for a powerful jet with $\epsilon_{\rm j} \sim 1$.

@@

\subsection{Black Hole Accretion Rates}
\label{sec:mdot}

@@

\subsection{Nuclear X-ray Luminosities}
\label{sec:Lx}

@@

\subsection{Local Growth Time of SMBHs }
\label{sec:growth}

@@

\subsection{TDE Jets}
\label{sec:TDE}

@@

\subsection{Regulation by Kinetic Outflow Feedback}
\label{sec:kinetic}

We focused on Compton radiative feedback in this paper, since its efficacy was straightforward to calculate in terms of the properties of the accretion flow.  However, if kinetic outflow feedback from the jet or accretion disk could is important, this could also have some influence on limiting the accretion rate, especially onto high mass black holes.  The effective wind heating velocity due to kinetic heating rate is proportional to the SMBH mass to a positive power, $v_{w}^{\bullet} \propto M_{\bullet}^{0.33(0.28)}$ for core(cusp) galaxies, respectively (eq.~[\ref{eq:vjet}]).  Coupled with the dependence of accretion rate on wind heating, $\dot{M}/\dot{M}_{\rm edd} \propto M_{\bullet}^{0.76(0.48)}v_{w}^{-3.8(-2.4)}$ (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]), SMBH heating of the formed we have adopted would lead to an Eddington ratio dependence $\dot{M}/\dot{M}_{\rm edd} \propto M_{\bullet}^{-0.5(-0.2)}$.  This prediction of an increasing Eddington ratio is in better with the observed dependence of $\langle L_{X}/L_{\rm edd} \rangle$ in elliptical galaxies (Fig.~\ref{fig:miller}), suggesting a possible explanation of stable accretion maintained by kinetic feedback as an alternative explanation for X-ray luminosities of elliptical galaxies.  On the other hand, our assumption that outflows heat the gas uniformly in volume is arbitrary and would need to more rigorously justified by numerical studies of how jets or disk outflows couple energy to their gaseous environment (Refs).

%\subsection{Jet Power Correlation}

%Accretion rates onto SMBHs can in principle be estimated directly by assuming Bondi accretion and using high resolution X-ray observations to estimate the gas temperature and density near the Bondi radius.
%This technique has led to the intriguing conclusion that the Bondi accretion power correlates with the AGN jet power (e.g.,
%\citealt{AllenDunn+:2006a}; \citealt{Russell+13}; \citealt{FujitaKawakatu+:2014a}), the latter estimated from the energy
%required to inflate the observed cavities within the X-ray emitting gas.  In practice is it almost never possible to resolve the Bondi radius, therefore requiring all such studies to extrapolate the observed gas temperature and density profiles to smaller radial scales.

@@

Our results could in principle also have implications for nuclear star clusters or globular clusters.  {\bf BDM :Do we really want to talk about this?  or is this an apology? I will let Aleksey fill in his ideas...}
